% !TeX root = ../main.tex

\chapter{Related Work}\label{chapter:related}

\section{Neural radiance fields}

\ac{NeRF}~\parencite{mildenhall2020nerf} introduced the idea of representing a
scene as a continuous volumetric function parameterized by a small MLP and
rendered by emission–absorption ray marching. Many follow-ups have
addressed its main limitations: training time
(InstantNGP~\parencite{mueller2022instant}), unbounded scenes
(Mip-NeRF~360), and editability.

\section{3D Gaussian Splatting}

\ac{3DGS}~\parencite{kerbl3Dgaussians} replaces the implicit MLP with an
explicit set of anisotropic 3D Gaussians, each carrying an opacity and
view-dependent \ac{SH} colour. A tile-based differentiable rasterizer
projects each Gaussian to screen space and composites them
back-to-front. The representation trains in minutes, renders in real time,
and serializes to a compact point cloud, which makes it attractive for
downstream tasks.

\section{Editing radiance fields}

Inserting or removing content in NeRF/3DGS scenes has been approached in
three broad ways:

\begin{description}
  \item[Latent editing.] Re-train or fine-tune the radiance field with
    additional supervision (e.g.\ text-guided diffusion) until the desired
    edit appears. Powerful but slow and unstable.
  \item[Hybrid representations.] Combine the radiance field with an
    explicit mesh or signed distance function so that classical mesh
    insertion tools work directly.
  \item[Compositing.] Treat the host scene as a passive background and
    render the inserted object separately, then composite. This is the
    route taken by InsertGS because it preserves the host 3DGS scene
    untouched.
\end{description}

\section{Image-based lighting}

Capturing real-world illumination as an environment map for shading virtual
objects dates back to Debevec~\parencite{debevec1998rendering}. The main
challenge in our setting is that the “environment” is not a chrome ball but
a learned 3DGS reconstruction; we therefore need a way to rasterize the
Gaussian cloud into a cubemap that is geometrically consistent with the
captured room.
